---
output:
  word_document: default
  html_document: default
---
# Module 4 - Assignment 1
## Amanda Odum

```{r message = "FALSE"}
options(tidyverse.quiet = TRUE)
library(tidyverse)
library(tidymodels)
library(caret)
library(rpart) 
library(rpart.plot) 
library(RColorBrewer) 
library(rattle)  
```

```{r data 1}
heart <- read_csv("heart_disease-1.csv")
str(heart)
summary(heart)
```

```{r factor conversion 1}
heart = heart %>% mutate(HeartDisease = as_factor(HeartDisease)) %>% 
  mutate(HeartDisease = fct_recode(HeartDisease, "No" = "0", "Yes" = "1" )) %>%
  mutate(ChestPainType = as_factor(ChestPainType)) %>% mutate(Sex = as_factor(Sex)) %>%
  mutate(RestingECG = as_factor(RestingECG)) %>% 
  mutate(ExerciseAngina = as_factor(ExerciseAngina))

str(heart)
```
**Task 1**
```{r training and testing}
set.seed(12345) 
heart_split = initial_split(heart, prop = 0.7, strata = HeartDisease) #70% in training
train = training(heart_split) 
test = testing(heart_split)
```

**Task2**
```{r class tree 1}
heart_recipe = recipe(HeartDisease ~., train)

tree_model = decision_tree() %>% 
  set_engine("rpart", model = TRUE) %>%
  set_mode("classification")

heart_wflow = 
  workflow() %>% 
  add_model(tree_model) %>% 
  add_recipe(heart_recipe)

heart_fit = fit(heart_wflow, train)

heart_fit %>%
  pull_workflow_fit() %>%
  pluck("fit")  

tree = heart_fit %>% #extract the tree's fit from the fit object
  pull_workflow_fit() %>% 
  pluck("fit")

rpart.plot(tree) #plot the tree
```
**Task 3**
```{r examine cp}
heart_fit$fit$fit$fit$cptable #examine cp values
```

*Two splits (CP=0.01742160) was the optimized cp value tried by R as this resulted in the lowest xerror value of 0.3588850.*

**Task 4**
```{r tuning grid 1 }
set.seed(123)
folds = vfold_cv(train, v = 5)

heart_recipe = recipe(HeartDisease ~., train) %>%
  step_dummy(all_nominal(),-all_outcomes())

tree_model = decision_tree(cost_complexity = tune()) %>% 
  set_engine("rpart", model = TRUE) %>% 
  set_mode("classification")

tree_grid = grid_regular(cost_complexity(),
                          levels = 25)
heart_wflow = 
  workflow() %>% 
  add_model(tree_model) %>% 
  add_recipe(heart_recipe)

tree_res = 
  heart_wflow %>% 
  tune_grid(
    resamples = folds,
    grid = tree_grid
    )

tree_res %>%
  collect_metrics() %>%
  ggplot(aes(cost_complexity, mean)) +
  geom_line(size = 1.5, alpha = 0.6) +
  geom_point(size = 2) +
  facet_wrap(~ .metric, scales = "free", nrow = 2) 
```

**Task 5**
```{r best accuracy 1}
best_tree = tree_res %>%
  select_best("accuracy")

best_tree
```

**Task 6**
```{r best tree 1 }
final_wf = 
  heart_wflow %>% 
  finalize_workflow(best_tree)

final_fit = fit(final_wf, train)

tree = final_fit %>% 
  pull_workflow_fit() %>% 
  pluck("fit")

fancyRpartPlot(tree, tweak = 1.5)
```

**Task 7**
```{r accuracy tree 1}
treepred1 = predict(final_fit, train, type = "class")
head(treepred1)

confusionMatrix(treepred1$.pred_class,train$HeartDisease,positive="Yes")

treepred1_test = predict(final_fit, test, type = "class")
head(treepred1_test)

confusionMatrix(treepred1_test$.pred_class,test$HeartDisease,positive="Yes")
```
*The accuracy is 82.61% on the test data set.*

**Task 8**
```{r data 2}
blood <- read_csv("Blood.csv")

blood = blood %>% mutate(DonatedMarch = as_factor(DonatedMarch)) %>% 
  mutate(DonatedMarch = fct_recode(DonatedMarch, "No" = "0", "Yes" = "1" ))
```

**Task 9**
```{r training and testing 2}
set.seed(1234) 
blood_split = initial_split(blood, prop = 0.7, strata = DonatedMarch) #70% in training
train2 = training(blood_split) 
test2 = testing(blood_split)
```

```{r class tree 2}
blood_recipe = recipe(DonatedMarch ~., train2)

tree_model = decision_tree() %>% 
  set_engine("rpart", model = TRUE) %>%
  set_mode("classification")

blood_wflow = 
  workflow() %>% 
  add_model(tree_model) %>% 
  add_recipe(blood_recipe)

blood_fit = fit(blood_wflow, train2)

blood_fit %>%
  pull_workflow_fit() %>%
  pluck("fit")  

tree2 = blood_fit %>% #extract the tree's fit from the fit object
  pull_workflow_fit() %>% 
  pluck("fit")

rpart.plot(tree2) #plot the tree
```
```{r tuning grid 2}
set.seed(1234)
folds = vfold_cv(train2, v = 5)

blood_recipe = recipe(DonatedMarch ~., train2) %>%
  step_dummy(all_nominal(),-all_outcomes())

tree_model2 = decision_tree(cost_complexity = tune()) %>% 
  set_engine("rpart", model = TRUE) %>% 
  set_mode("classification")

tree_grid2 = grid_regular(cost_complexity(),
                          levels = 25)
blood_wflow = 
  workflow() %>% 
  add_model(tree_model2) %>% 
  add_recipe(blood_recipe)

tree_res2 = 
  blood_wflow %>% 
  tune_grid(
    resamples = folds,
    grid = tree_grid2
    )

tree_res2 %>%
  collect_metrics() %>%
  ggplot(aes(cost_complexity, mean)) +
  geom_line(size = 1.5, alpha = 0.6) +
  geom_point(size = 2) +
  facet_wrap(~ .metric, scales = "free", nrow = 2) 
```
```{r best accuracy 2}
best_tree2 = tree_res2 %>%
  select_best("accuracy")

best_tree2
```
*0.01778279 appears to maximize accuracy.*

**Task 10**
```{r best tree 2}
final_wf = 
  blood_wflow %>% 
  finalize_workflow(best_tree2)

final_fit = fit(final_wf, train2)

tree2 = final_fit %>% 
  pull_workflow_fit() %>% 
  pluck("fit")

fancyRpartPlot(tree2, tweak = 1.5)
```

**Task 11**
```{r accuracy tree 2}
treepred2 = predict(final_fit, train2, type = "class")
head(treepred2)

confusionMatrix(treepred2$.pred_class,train2$DonatedMarch,positive="Yes")

treepred2_test = predict(final_fit, test2, type = "class")
head(treepred2_test)

confusionMatrix(treepred2_test$.pred_class,test2$DonatedMarch,positive="Yes")
```
*Accuracy on the training set is 80.69% and on the testing set is 78.22%. The tree performed well on both sets, though with higher accuracy on the training set.* 
