---
output:
  word_document: default
  html_document: default
---
# Module 4 - Assignment 2
## Amanda Odum

```{r include = FALSE}
library(tidyverse)
library(tidymodels)
library(gridExtra) 
library(vip)
library(ranger) #for random forests
library(caret)
library(VIM)
library(skimr)
```

```{r data}
drug <- read_csv("drug_data-1.csv")
names(drug) = c("ID", "Age", "Gender", "Education", "Country", "Ethnicity",
"Nscore", "Escore", "Oscore", "Ascore", "Cscore", "Impulsive",
"SS", "Alcohol", "Amphet", "Amyl", "Benzos", "Caff", "Cannabis",
"Choc", "Coke", "Crack", "Ecstasy", "Heroin", "Ketamine", "Legalh",
"LSD", "Meth", "Mushrooms", "Nicotine", "Semer", "VSA")
#str(drug)

drug[drug == "CL0"] = "No"
drug[drug == "CL1"] = "No"
drug[drug == "CL2"] = "Yes"
drug[drug == "CL3"] = "Yes"
drug[drug == "CL4"] = "Yes"
drug[drug == "CL5"] = "Yes"
drug[drug == "CL6"] = "Yes"

drug_clean = drug %>% mutate_at(vars(Age:Ethnicity), funs(as_factor)) %>%
mutate(Age = factor(Age, labels = c("18_24", "25_34", "35_44",
"45_54", "55_64", "65_"))) %>%
mutate(Gender = factor(Gender, labels = c("Male", "Female"))) %>%
mutate(Education = factor(Education, labels =
c("Under16", "At16", "At17", "At18", "SomeCollege",
"ProfessionalCert", "Bachelors", "Masters", "Doctorate"))) %>%
mutate(Country = factor(Country,
labels = c("USA", "NewZealand", "Other", "Australia",
"Ireland","Canada","UK"))) %>%
mutate(Ethnicity = factor(Ethnicity,
labels = c("Black", "Asian", "White", "White/Black", "Other",
"White/Asian", "Black/Asian"))) %>% mutate_at(vars(Alcohol:VSA), funs(as_factor)) %>%
select(-ID)

#str(drug_clean)

drug_clean = drug_clean %>% select(!(Alcohol:Mushrooms)) %>% select(!(Semer:VSA))
names(drug_clean)

#summary(drug_clean)

```

**Task 1**
```{r check for missing data}
skim(drug_clean)
```
*No data is missing.*

**Task 2**
```{r training and testing}
set.seed(1234) 
drug_clean_split = initial_split(drug_clean, prop = 0.7, strata = Nicotine) #70% in training
train = training(drug_clean_split)
test = testing(drug_clean_split)
```

**Task 3**
```{r visual1}
p1 = ggplot(train, aes(x = Age, fill = Nicotine)) + geom_bar(position = "fill") + theme(axis.text.x = element_text(angle = 60, hjust = 1))
p2 = ggplot(train, aes(x = Gender, fill = Nicotine)) + geom_bar(position = "fill") + theme(axis.text.x = element_text(angle = 60, hjust = 1))
p3 = ggplot(train, aes(x = Education, fill = Nicotine)) + geom_bar(position = "fill") + theme(axis.text.x = element_text(angle = 60, hjust = 1))
p4 = ggplot(train, aes(x = Country, fill = Nicotine)) + geom_bar(position = "fill") + theme(axis.text.x = element_text(angle = 60, hjust = 1))
grid.arrange(p1,p2,p3,p4)
```
```{r visual 2}
p5 = ggplot(train, aes(x = Ethnicity, fill = Nicotine)) + geom_bar(position = "fill")+ theme(axis.text.x = element_text(angle = 60, hjust = 1))
p6 = ggplot(train, aes(x = Nicotine, y = Nscore)) + geom_boxplot()
p7 = ggplot(train, aes(x = Nicotine, y = Escore)) + geom_boxplot()
p8 = ggplot(train, aes(x = Nicotine, y = Oscore)) + geom_boxplot()
grid.arrange(p5, p6, p7, p8)
```

```{r visual 3}
p9 = ggplot(train, aes(x = Nicotine, y = Ascore)) + geom_boxplot()
p10 = ggplot(train, aes(x = Nicotine, y = Cscore)) + geom_boxplot()
p11 = ggplot(train, aes(x = Nicotine, y = Impulsive)) + geom_boxplot()
p12 = ggplot(train, aes(x = Nicotine, y = SS)) + geom_boxplot()
grid.arrange(p9, p10, p11, p12)
```
*Nicotine usage appears to drop continually as age increases.  Females appear ~1.5 times less likely to use nicotine than males. Nicotine usage appears to be negatively correlated with education levels, as it goes down as education level increases.  Nicotine usage is highest in the US and Ireland, and lowest in the UK and New Zealand.  People who identify as black and Asian appear less likely to use nicotine than other races.  Nicotine users have higher Nscore, Oscore, Impulsive,and SS values.  Nicotine users have lower Ascore and Cscore values, and have roughtly the same Escore values as those who do not use nicotine.*

**Task 4**
```{r random forest}
set.seed(123)
rf_folds = vfold_cv(train, v = 5)

drug_recipe = recipe(Nicotine ~., train) %>%
  step_dummy(all_nominal(), -all_outcomes())

rf_model = rand_forest(mtry = tune(), min_n = tune(), trees = 100) %>% 
  set_engine("ranger", importance = "permutation") %>% #added importance metric
  set_mode("classification")

drug_wflow = 
  workflow() %>% 
  add_model(rf_model) %>% 
  add_recipe(drug_recipe)

rf_grid = grid_regular(
  mtry(range = c(2, 8)), 
  min_n(range = c(5, 20)), 
  levels = 10
)

set.seed(123)
rf_res_tuned = tune_grid(
  drug_wflow,
  resamples = rf_folds,
  grid = rf_grid #use the tuning grid
)
```

```{r parameter visual}
rf_res_tuned %>%
  collect_metrics() %>%
  filter(.metric == "accuracy") %>%
  select(mean, min_n, mtry) %>%
  pivot_longer(min_n:mtry,
    values_to = "value",
    names_to = "parameter"
  ) %>%
  ggplot(aes(value, mean, color = parameter)) +
  geom_point(show.legend = FALSE) +
  facet_wrap(~parameter, scales = "free_x") +
  labs(x = NULL, y = "Accuracy")

rf_res_tuned %>%
  collect_metrics() %>%
  filter(.metric == "accuracy") %>%
  mutate(min_n = factor(min_n)) %>%
  ggplot(aes(mtry, mean, color = min_n)) +
  geom_line(alpha = 0.5, size = 1.5) +
  geom_point() +
  labs(y = "Accuracy")
```

**Task 5**
```{r best mtry and min_n}
best_rf = select_best(rf_res_tuned, "accuracy")

final_rf = finalize_workflow(
  drug_wflow,
  best_rf
)

final_rf
final_rf_fit = fit(final_rf, train)
```

**Task 6**
```{r fit for training}

trainpredrf = predict(final_rf_fit, train)
head(trainpredrf)

confusionMatrix(trainpredrf$.pred_class, train$Nicotine, positive = "Yes")
```
```{r fit for testing}
testpredrf = predict(final_rf_fit, test)
head(testpredrf)

confusionMatrix(testpredrf$.pred_class, test$Nicotine, positive = "Yes")
```

*The model performs much better on the training set than on the testing set.*

**Task 7**

*This model could be used by insurance companies (life or health) to deny coverage and/or increase premiums or by a clinician or healthcare organization to make healthcare decisions on things such as clinical trial or organ transplant on an individual.  I would not recommend this model for real world use as its accuracy on the testing set was only 61.5%. My concern would be that this could incorrectly classify an individual as a smoker/nicotine user and prevent them from receiving appropriate care or coverage based on factors such as age, location, education, et cetera.*  